---
title: "Riparian Invertebrates: Data Import"
output: html_notebook
---

##Packages
```{r}
# packages
library(tidyverse)
library(readxl)
library(forcats)
```

##Import data 
```{r}
#importing all as text to begin with
ers_data <- read_excel("C:/Christoph/riparianInverts/data/RIPARIA 27March2015.xlsx", sheet = 1, col_types = "text")

BRC_carabid_data <- read_excel("C:/Christoph/riparianInverts/data/RIPARIA 27March2015.xlsx", sheet = 2, col_types = "text")

isr_data <- read_excel("C:/Christoph/riparianInverts/data/RIPARIA 27March2015.xlsx", sheet = 3, col_types = "text")
```


##ERS Data

###Inspect

Start with abundance data, its a quite a mix
```{r}
length(unique(ers_data$Abundance_Data))
```

#### Get a handle on numeric data

Try to find all the numeric values
```{r}
#new column, convert abundance to numeric. 
ers_data$abund_num <- as.numeric(ers_data$Abundance_Data)
#all data containing a non-numeric is now NA, all numeric is numeric
plot(ers_data$abund_num)
```

Looks like there's a big gap in the middle where there is either non numeric or no data. Also looks like most of the data is well below 500.  

```{r}
abund_num <- ers_data %>% 
  select(abund_num) %>%  
  filter(is.na(abund_num) == FALSE) 
abund_num <- abund_num$abund_num

length(unique(abund_num))
length(abund_num)
```
260 unique numeric values.  8555 records with numeric abundance values


Letzs zoom in: 
```{r}
hist(abund_num, breaks = 10000, xlim = c(0,100))
```


```{r}
ers_hist <- hist(abund_num, breaks = 100, plot = FALSE)
knitr::kable(tibble(ers_hist$breaks[-1], ers_hist$counts))
```

The majority of numeric abundance records are below 50 (8073).  A few values in the thousands and four-thousands.  

#### Get a handle on non-numeric data

```{r}
ers_data <- add_column(ers_data, abund_char = ers_data$Abundance_Data) 
  
ers_data$abund_char <- replace(ers_data$abund_char, 
                               is.na(ers_data$abund_num)==FALSE, 
                               NA)
  
  
```




```{r}
length(unique(ers_data$abund_char))
length(ers_data$abund_char[which(is.na(ers_data$abund_char)==FALSE)])
```
8051 observations; 136 unique abundance values


```{r}
ers_char <- data.frame(table(ers_data$abund_char))
ers_char <- arrange(ers_char, -Freq)
knitr::kable(ers_char)
```


what about sample types
```{r}
knitr::kable(
  ers_data %>%
    count(Sample_Type) %>%
    arrange(-n)
)
```


```{r}
wordcloud2::wordcloud2(table(ers_data$Sample_Type))
```

does data type relate to sample type
```{r}
sample_data_types <-
  ers_data %>%
    group_by(Sample_Type) %>%
    count("abund_character" = is.na(abund_char)==FALSE, 
        "abund_number" = is.na(abund_num)==FALSE) %>% 
    filter(abund_character == TRUE | abund_number == TRUE) 

sample_data_types$abund_character[which(sample_data_types$abund_character==TRUE)] <- 
  sample_data_types$n[which(sample_data_types$abund_character==TRUE)]
  
sample_data_types$abund_number[which(sample_data_types$abund_number==TRUE)] <- 
  sample_data_types$n[which(sample_data_types$abund_number==TRUE)]
  
sample_data_types$n <- NULL

knitr::kable(sample_data_types)
```


### Species

```{r}
wordcloud2::wordcloud2(table(ers_data$Species_name))
```

####River

```{r}
wordcloud2::wordcloud2(table(ers_data$`Sample River`))
```

####

```{r}
wordcloud2::wordcloud2(table(ers_data$Taxonomic_Family), size = 10)
```


```{r}
wordcloud2::wordcloud2(table(ers_data$`Author/Recorder`), size = 10)
```




