---
title: "Call notes 19 August 2020 "
output: html_notebook
---



```{r setup, include=FALSE}
# document rendering options
knitr::opts_chunk$set(
	message = FALSE,
	warning = FALSE,
	include = TRUE, 
	echo = FALSE,
	dpi = 300
)

# load packages
library(tidyverse)
library(cowplot)
library(iNEXT)
library(vegan)
library(VennDiagram)
library(grid)
library(gridExtra)


# generic graph settings
theme_doc <-  theme(axis.title = element_text(size = 10), 
                    text=element_text(size=10), 
                    plot.title = element_text(size = 11),  
                    legend.key.width = unit(1.2,"cm")) 

```


```{r}
#Import and prepare data
#The following script imports data, cleans it and prepares a number of datasets.  It also carries out a few analyses: 

source("../scripts/ripInverts-paper-code_2.R", chdir = T)
```


### data issues

ck to do a data dictionary that explains each table and to check number of sites / events / samples used in each analysis


### Effort

```{r}
effort_all <- 
  obs_all %>% select(spp_name, event_code) %>% 
  left_join(select(effort_data, event_code, effort)) %>% 
  left_join(select(spp_lookup_fam, species, 'Wetland species'), 
            by = c("spp_name" = "species")) %>% 
  rename(wetland_spp = 'Wetland species') %>% 
  unique() %>% 
  group_by(event_code, effort) %>% 
  count(wetland_spp) %>% 
  #filter(wetland_spp == 1) %>% 
  filter(event_code != "unknown")

effort_all %>% arrange(-n)

plot_effort <- ggplot(effort_all, 
                              aes(x = as.factor(effort), y = n), 
                              fill = as.factor(effort))+
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  #geom_violin(scale = "width")+
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  labs(x = "Sampling effort", y = "Number of species", title = "B") +
  theme_bw() +
  theme_doc
  
plot_effort
```

```{r}
plot_effort_regr <- ggplot(effort_all, 
                              aes(x = effort, y = n), 
                              fill = as.factor(effort))+
  geom_boxplot(varwidth = T, colour = "darkgrey", aes(group = effort)) +
  geom_smooth(method = "lm")+
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  scale_x_continuous(breaks = 2:13) +
  labs(x = "Sampling effort", y = "Number of species", title = "B") +
  theme_bw() + theme_doc; plot_effort_regr

lm.eff <- lm(effort ~ n, effort_all)
broom::tidy(lm.eff)
summary(lm.eff)
lm.eff


lmer.eff <- lme4::lmer(effort ~ n + (1 | wetland_spp), data = effort_all)
summary(lmer.eff)
```







### Permanova

```{r}
##Dissimilarity----
#calculate jaccard dissimilarity
adon1 <- vegan::adonis2(selected_data ~ river + event_code + sample_type, data = selected_env, permutations = 120, method = "jaccard", by = "terms")

#test assumptions
dist_selected_data <- vegdist(x = selected_data, method = "jaccard")

bd.river <- betadisper(d = dist_selected_data, group = selected_env$river)
boxplot(bd.river)
# F-Test
anova(bd.river)
permutest(bd.river) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met


bd.event <- betadisper(d = dist_selected_data, group = selected_env$event_code)
boxplot(bd.event)
# F-Test
anova(bd.event)
permutest(bd.event) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met


bd.sample_type <- betadisper(d = dist_selected_data, group = selected_env$sample_type)
boxplot(bd.sample_type)
# F-Test
anova(bd.sample_type)
permutest(bd.sample_type) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met
```

```{r}
#re-do beta dispersion with a 'bray' distance matrix

#test assumptions
dist_selected_data <- vegdist(x = selected_data^0.25, method = "bray") #note: cube root transformed

bd.river <- betadisper(d = dist_selected_data, group = selected_env$river)
boxplot(bd.river)
# F-Test
anova(bd.river)
permutest(bd.river) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met


bd.event <- betadisper(d = dist_selected_data, group = selected_env$event_code)
boxplot(bd.event)
# F-Test
anova(bd.event)
permutest(bd.event) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met


bd.sample_type <- betadisper(d = dist_selected_data, group = selected_env$sample_type)
boxplot(bd.sample_type)
# F-Test
anova(bd.sample_type)
permutest(bd.sample_type) 
# We can/cannot find a statistically different dispersion 
# -> assumption of homogeneity is [not] met
```


```{r}
#do we have an equal number of hs and pf
selected_env %>% count(sample_type)


```

## Lets ordinate

```{r}
mds <- metaMDS(comm = selected_data^2, distance = "euclidian")

ordiplot(ord = mds, display = "species")
ordihull(ord = mds, groups = selected_env$sample_type, label = T)
```

#### conclusion of this bit

* report that we've tested beta dispersion and have found non-homog of dispersion (report the results for the three factors)

* report the results of the permanova with above proviso



### try restricting permutation scheme

```{r}
##Dissimilarity----
#calculate jaccard dissimilarity
adon1 <- vegan::adonis(selected_data ~ event_code + sample_type, data = selected_env, strata = selected_env$river, permutations = 120, method = "jaccard", by = "terms")
```

 makes no difference






































