---
title: "Riparian Invertebrates: Accumulation curves"
output: html_notebook
---

##Packages
```{r}
# packages
library(tidyverse)
library(readxl)
library(forcats)
```

##Import data
```{r}
effort_data <- read_csv("../data/effort.csv") 
field_data <- read_csv("../data/field_data_selected.csv")
field_data <-  field_data %>% 
  #get rid of event = NA
  filter(!is.na(event))

indices <- read_csv("../data/indices.csv")
river_indices <- read_csv("../data/river_indices.csv")
```

## Plot

```{r}
ggplot(indices %>%  
         gather(key = index, value = score, n_wetland:sqs_marsh, na.rm = T), 
       aes(x = effort, y = score))+
  geom_point() +
  geom_smooth(method = "lm")+
  scale_x_continuous(breaks = seq(1:12)) +  
  facet_wrap(~index)
```


```{r}
ggplot(river_indices %>%  
         gather(key = index, value = score, n_wetland:sqs_marsh, na.rm = T) %>% 
         filter(river %in% c("Severn", "Towy", "Wye")) %>%
         filter(index %in% c("sqi_marsh", "sqi_runWater")) %>% 
         droplevels(), 
       aes(x = effort, y = score))+
  geom_point() +
  geom_smooth(method = "lm")+
  scale_x_continuous(breaks = seq(1:12)) +  
  theme(panel.grid.minor = element_blank()) +
  facet_grid(river~index)
```


```{r}
ggplot(river_indices %>%  
         gather(key = index, value = score, n_wetland:sqs_marsh, na.rm = T) %>% 
         filter(river %in% c("Severn", "Towy", "Wye")) %>%
         filter(index %in% c("sqi_marsh", "sqi_runWater")) %>% 
         droplevels(), 
       aes(x = effort, y = score))+
  geom_point(position = position_jitter(width = 0.5, height = 20)) +
  #geom_point(position = "jitter")+
  geom_smooth(method = "lm")+
  scale_x_continuous(breaks = seq(1:12)) +  
  theme(panel.grid = element_blank()) +
  facet_grid(river~index)
```


```{r}
library(vegan)

rarecurve()
```




```{r}
table(field_data$abund_char)
table(field_data$abund_num)
table(field_data$abund_char, field_data$sample_type)
```

```{r}
field_data %>% 
  filter(event == "Monnow 2015") %>% 
  count(spp_name) %>% 
  arrange(-n)
```


Lets convert this to a frequency matrix
```{r}
field_data_freq <- field_data %>% 
  #get rid of event = NA
  filter(!is.na(event)) %>% 
  #calculate frequencies
  group_by(river, event, spp_name) %>% 
  summarise(n = length(spp_name)) %>% 
  #abbreviate species names
  mutate(spp_name = vegan::make.cepnames(spp_name)) %>% 
  #spread to a matrix
  spread(key = spp_name, value = n, fill = 0) %>%
  data.frame() %>%
  arrange(river) 

field_data_freq


field_data_pitfall <- field_data %>% 
  #get rid of event = NA
  filter(!is.na(event)) %>% 
  #select by sample type 
  filter(sample_type == "pitfall") %>% 
  #calculate frequencies
  group_by(river, event, spp_name) %>% 
  summarise(n = length(spp_name)) %>% 
  #abbreviate species names
  mutate(spp_name = vegan::make.cepnames(spp_name)) %>% 
  #spread to a matrix
  spread(key = spp_name, value = n, fill = 0) %>%
  data.frame() %>%
  arrange(river) 


field_data_search <- field_data %>% 
  #get rid of event = NA
  filter(!is.na(event)) %>% 
  #select by sample type 
  filter(sample_type != "pitfall") %>% 
  #calculate frequencies
  group_by(river, event, spp_name) %>% 
  summarise(n = length(spp_name)) %>% 
  #abbreviate species names
  mutate(spp_name = vegan::make.cepnames(spp_name)) %>% 
  #spread to a matrix
  spread(key = spp_name, value = n, fill = 0) %>%
  data.frame() %>%
  arrange(river) 


```

species accumulation curve for all surveys
```{r}
spec_accum <- vegan::specaccum(comm = field_data_freq)

plot(spec_accum)
```

species accumulation curve sy sampple type
```{r}
spec_accum_pit <- vegan::specaccum(comm = select(field_data_pitfall, -c(river, event)))
spec_accum_search <- specaccum(comm = select(field_data_search, -c(river, event)))

plot(spec_accum)
plot(spec_accum_search, col = "green", add = TRUE)
plot(spec_accum_pit, col = "red", add = TRUE)

```



species accumulation curve for selected surveys
```{r}
library(vegan)

#list of surveys on selected river
events_severn <- unique(field_data$event[which(field_data$river == "Severn")])
#species accumulation curve by resampling events
spec_accum_severn <- specaccum(comm = field_data_freq[events_severn,])
plot(spec_accum_severn, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Severn")
abline(h = max(spec_accum_severn$richness), col = "black")
abline(h = max(spec_accum_severn$richness)*0.95, col = "red")
abline(h = max(spec_accum_severn$richness)*0.9, col = "red", lty = "dotted")

#list of surveys on selected river
events_towy <- unique(field_data$event[which(field_data$river == "Towy")])
#species accumulation curve by resampling events
spec_accum_towy <- specaccum(comm = field_data_freq[events_towy,])
plot(spec_accum_towy, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Towy")
abline(h = max(spec_accum_towy$richness), col = "black")
abline(h = max(spec_accum_towy$richness)*0.95, col = "red")
abline(h = max(spec_accum_towy$richness)*0.9, col = "red", lty = "dotted")

#list of surveys on selected river
events_wye <- unique(field_data$event[which(field_data$river == "Wye")])
#species accumulation curve by resampling events
spec_accum_wye <- specaccum(comm = field_data_freq[events_wye,])
plot(spec_accum_wye, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Wye")
abline(h = max(spec_accum_wye$richness), col = "black")
abline(h = max(spec_accum_wye$richness)*0.95, col = "red")
abline(h = max(spec_accum_wye$richness)*0.9, col = "red", lty = "dotted")


plot(spec_accum_wye, col = "green")
plot(spec_accum_severn, add = T, col = "blue")
plot(spec_accum_towy, add = T, col = "orange")

# op <- par()
# par(mfrow = c(1,3))
# plot(spec_accum)
# plot(spec_accum_severn, col = "red", add = T)
#par(op)
```

Try plotting specaccum object in ggplot
```{r}
graphdata <- data.frame(n_sites = spec_accum$sites,
                        n_spp = spec_accum$richness,
                        sd = spec_accum$sd)

ggplot(graphdata, aes(x = n_sites, y = n_spp))+
  geom_point() +
  geom_errorbar(aes(ymin = n_spp-sd, ymax = n_spp+sd))

ggplot(graphdata, aes(x = n_sites, y = n_spp))+
  geom_point() +
  geom_path(aes(y = n_spp-sd)) +
  geom_path(aes(y = n_spp+sd))

plot(spec_accum)
```














Example below using rarefaction function in package vegan.  It doesn't make sense because it is intended for individual based sampling rather than sample based sampling.  
```{r}
S <- specnumber(field_data_freq) # observed number of species
raremax <- min(rowSums(field_data_freq))
Srare <- rarefy(field_data_freq, raremax)
plot(S, Srare, xlab = "Observed No. of Species", ylab = "Rarefied No. of Species")
abline(0, 1)
rarecurve(field_data_freq, step = 20, sample = raremax, col = "blue", cex = 0.6)
```


better to use package iNext but this isn't avail for R version 3.3.2
```{r}
#install.packages("iNEXT")
#need version 3.4!!!
```




Frequency matrix for selected rivers
```{r}
field_data_freq <- field_data %>% 
  #get rid of event = NA
  filter(!is.na(event)) %>% 
  #calculate frequencies
  group_by(event, spp_name) %>% 
  summarise(n = length(spp_name)) %>% 
  #abbreviate species names
  mutate(spp_name = vegan::make.cepnames(spp_name)) %>% 
  #spread to a matrix
  spread(key = spp_name, value = n, fill = 0) %>%
  data.frame() %>% 
  column_to_rownames(var = "event")

field_data_freq

```