---
title: "Riparian Inverts: ordination"
output: html_notebook
---

```{r}
library(tidyverse)
library(vegan)
#library(MASS)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
source("../R/ripInverts-paper-code_1.R", chdir = T)
```

##Ordination - all records

```{r}
#split matrix into species data and environmental data
spp_dat <- select(obs_all_mat, -c(river:sample_type))

env_dat <- select(obs_all_mat, c(river:sample_type))
```


```{r}
#calculate distance matrix 
spp_dist1 <- vegdist(spp_dat)

#ordinate with NMDS using Bray-Curtis dissimilarity
spp_mds_bray <- metaMDS(spp_dat, trace = FALSE)
spp_mds_bray
```
**commentary below adapted from Oksanen, J. http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf **
The species data was not tranformed.  Distance was calculated using Bray-Curtis dissimilarity.  Function run isoMDS with several random starts, and stopped either after a certain number of tries, or after finding two similar configurations with minimum stress. In any case, it returned the
best solution.  Function rotated the solution so that the largest variance of site
scores will be on the first axis. Function scaled the solution so that one unit corresponds to halving
of community similarity from the replicate similarity.   Function found species scores as weighted averages of site scores, but expanded them so that species and site scores have equal variances. This expansion can be undone using 'shrink = TRUE' in display commands.


```{r}
#ordinate with NMDS using Jaccard dissimilarity
spp_mds_jac <- metaMDS(spp_dat, trace = FALSE, distance = "jac")
spp_mds_jac
```

```{r}
#compare jaccard v bray NMDS

procrustes(spp_mds_bray, spp_mds_jac)
plot(procrustes(spp_mds_bray, spp_mds_jac))
plot(procrustes(spp_mds_bray, spp_mds_jac), kind = 2)
```


```{r}
plot(spp_mds_bray)
plot(spp_mds_jac)

spp_mds_data <- cbind(env_dat, spp_mds_jac$points)
hist(spp_mds_data$MDS1)

ggplot(filter(spp_mds_data, MDS1 < 0), aes(MDS1, MDS2, color = sample_type)) + geom_point() + theme_bw()

ggplot(filter(spp_mds_data, MDS1 < 0), aes(MDS1, MDS2, color = river)) + geom_point() + theme_bw()

```
REMOVED OUTLIERS FROM RESULTS: PROBABLY BETTER TO REMOVE FROM DATA

```{r}
#ordinate with PCA
spp_pca <- rda(spp_dat, scale = TRUE)
spp_pca  
plot(spp_pca, scaling = -3)  
biplot(spp_pca, scaling = 2)
```

```{r}
#ordinate with correspondence analysis
spp_ca <- cca(spp_dat)
spp_ca

plot(spp_ca)

```


```{r}
#ordinate with detrended correspondence analysis
spp_dca <- decorana(spp_dat)

plot(spp_dca, display = "sites", type = "points")
with(env_dat, ordiellipse(spp_dca, sample_type, col = "blue", kind = "se", conf = 0.95))
with(env_dat, ordispider(spp_dca, sample_type, col = "blue", label= TRUE))
with(env_dat, ordihull(spp_dca, sample_type, col="blue", lty=2))


plot(spp_dca, display = "sites", type = "points")
with(env_dat, ordiellipse(spp_dca, river, kind = "se", conf = 0.95))
with(env_dat, ordispider(spp_dca, river, col = "blue", label= TRUE))
with(env_dat, ordihull(spp_dca, river, col="blue", lty=2))

plot(spp_dca, display = "sites", type = "points")
with(env_dat, ordiellipse(spp_dca, location, kind = "se", conf = 0.95))
with(env_dat, ordispider(spp_dca, location, col = "blue", label= TRUE))
with(env_dat, ordihull(spp_dca, location, col="blue", lty=2))
```

```{r}
spp_dca_data <- cbind(env_dat, spp_dca$rproj)

ggplot(spp_dca_data, aes(DCA4, DCA2, color = sample_type)) + geom_point() + theme_bw()

ggplot(spp_dca_data, aes(DCA4, DCA2, color = river)) + geom_point() + theme_bw()

```


##Ordination - recent records

```{r}
#split matrix into species data and environmental data
spp_rct_dat <- select(obs_rct_mat, -c(river:sample_type))

env_rct_dat <- select(obs_rct_mat, c(river:sample_type))
```


```{r}
#calculate distance matrix 
spp_rct_dist1 <- vegdist(spp_rct_dat)

#ordinate with NMDS using Bray-Curtis dissimilarity
spp_rct_mds_bray <- metaMDS(spp_rct_dat, trace = FALSE)
spp_rct_mds_bray
```
**commentary below adapted from Oksanen, J. http://cc.oulu.fi/~jarioksa/opetus/metodi/vegantutor.pdf **
The species data was not tranformed.  Distance was calculated using Bray-Curtis dissimilarity.  Function run isoMDS with several random starts, and stopped either after a certain number of tries, or after finding two similar configurations with minimum stress. In any case, it returned the
best solution.  Function rotated the solution so that the largest variance of site
scores will be on the first axis. Function scaled the solution so that one unit corresponds to halving
of community similarity from the replicate similarity.   Function found species scores as weighted averages of site scores, but expanded them so that species and site scores have equal variances. This expansion can be undone using 'shrink = TRUE' in display commands.


```{r}
#ordinate with NMDS using Jaccard dissimilarity
spp_rct_mds_jac <- metaMDS(spp_rct_dat, trace = FALSE, distance = "jac")
spp_rct_mds_jac
```

```{r}
#compare jaccard v bray NMDS

procrustes(spp_rct_mds_bray, spp_rct_mds_jac)
plot(procrustes(spp_rct_mds_bray, spp_rct_mds_jac))
plot(procrustes(spp_rct_mds_bray, spp_rct_mds_jac), kind = 2)
```


```{r}
plot(spp_rct_mds_bray)
plot(spp_rct_mds_jac)

spp_rct_mds_data <- cbind(env_rct_dat, spp_rct_mds_jac$points)

ggplot(spp_rct_mds_data, aes(MDS1, MDS2, color = sample_type)) + geom_point() + theme_bw()

ggplot(spp_rct_mds_data, aes(MDS1, MDS2, color = river)) + geom_point() + theme_bw()

```
REMOVED OUTLIERS FROM RESULTS: PROBABLY BETTER TO REMOVE FROM DATA

```{r}
#ordinate with PCA
spp_rct_pca <- rda(spp_rct_dat, scale = TRUE)
spp_rct_pca  
plot(spp_rct_pca)  
biplot(spp_rct_pca, scaling = 2)
```

```{r}
#ordinate with correspondence analysis
spp_rct_ca <- cca(spp_rct_dat)
spp_rct_ca

plot(spp_rct_ca)

```


```{r}
#ordinate with detrended correspondence analysis
spp_rct_dca <- decorana(spp_rct_dat)

plot(spp_rct_dca, display = "sites", type = "points")
with(env_rct_dat, ordiellipse(spp_rct_dca, sample_type, col = "blue", kind = "se", conf = 0.95))
with(env_rct_dat, ordispider(spp_rct_dca, sample_type, col = "blue", label= TRUE))
with(env_rct_dat, ordihull(spp_rct_dca, sample_type, col="blue", lty=2))


plot(spp_rct_dca, display = "sites", type = "points")
with(env_rct_dat, ordiellipse(spp_rct_dca, river, kind = "se", conf = 0.95))
with(env_rct_dat, ordispider(spp_rct_dca, river, col = "blue", label= TRUE))
with(env_rct_dat, ordihull(spp_rct_dca, river, col="blue", lty=2))

plot(spp_rct_dca, display = "sites", type = "points")
with(env_rct_dat, ordiellipse(spp_rct_dca, location, kind = "se", conf = 0.95))
with(env_rct_dat, ordispider(spp_rct_dca, location, col = "blue", label= TRUE))
with(env_rct_dat, ordihull(spp_rct_dca, location, col="blue", lty=2))
```

##ggplot

```{r}
#ordinate with detrended correspondence analysis
  #plottign in ggplot
spp_rct_dca <- decorana(spp_rct_dat)

spp_rct_dca_sites <- cbind(env_rct_dat, scores(spp_rct_dca, display = "sites", origin = TRUE))

spp_rct_dca_species <- scores(spp_rct_dca, display = "species", origin = TRUE) %>%
  data.frame() %>% 
  rownames_to_column("species")

#hulls
spp_rct_dca_sites %>% 
  filter(river == "Wye")

grp.a <- data.scores[data.scores$grp == "A", ][chull(data.scores[data.scores$grp == 
    "A", c("NMDS1", "NMDS2")]), ]  # hull values for grp A
grp.b <- data.scores[data.scores$grp == "B", ][chull(data.scores[data.scores$grp == 
    "B", c("NMDS1", "NMDS2")]), ]  # hull values for grp B

hull.data <- rbind(grp.a, grp.b)  #combine grp.a and grp.b
hull.data
```


```{r}
ggplot(spp_rct_dca_sites, aes(DCA1, DCA2, color = sample_type)) + 
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +
  coord_equal()+
  theme_bw()+ 
  labs(title = "Detrended correspondence analysis", subtitle = "by sample type")

ggplot(spp_rct_dca_sites, aes(DCA1, DCA2, color = river)) +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +
  coord_equal()+
  theme_bw()+ 
  labs(title = "Detrended correspondence analysis", subtitle = "by river")

ggplot(spp_rct_dca_sites, aes(DCA1, DCA2, colour = event_code)) +
  geom_point() + 
  facet_wrap(~river) +
  coord_equal()+
  theme_bw()+
  theme(legend.position = "none") +
  labs(title = "Detrended correspondence analysis", subtitle = "by river and event")
```

```{r}
ggplot() + 
  geom_point(data=spp_rct_dca_species,aes(x=DCA1,y=DCA2),alpha=0.5, size = 0.5, colour = "red") +  
  # add the species labels
  geom_point(data=spp_rct_dca_sites,aes(x=DCA1,y=DCA2,colour=sample_type), shape = 1) + # add the site markers
  #geom_text(data=spp_rct_dca_sites,aes(x=DCA1,y=DCA2, label=replicate),size=6,vjust=0) +  
  # add the site labels
  #scale_colour_manual(values=c("A" = "red", "B" = "blue")) +
  #coord_equal() +
  theme_bw()
```

```{r}
source("../extra_scripts/gg_ordiplot.R")
source("../extra_scripts/ord_labels.R")
```

```{r}
my.plot <- gg_ordiplot(spp_rct_dca, env_rct_dat$river, kind = "sd", 
            label = TRUE, spider = FALSE, hull = TRUE, ellipse = FALSE,
            plot = FALSE, pt.size = 2) 

my.plot$plot +
  theme_bw() +
  coord_equal() +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  geom_text(fontface = "bold") +
  labs(color="River", x="DCA 1", y="DCA 2", 
       title = "Detrended correspondence analysis", subtitle = "by river") +
  theme(legend.position = "none")
```

```{r}
my.plot <- gg_ordiplot(spp_rct_dca, env_rct_dat$sample_type, kind = "sd", 
            label = TRUE, spider = FALSE, hull = TRUE, ellipse = FALSE,
            plot = FALSE, pt.size = 2) 

my.plot$plot +
  theme_bw() +
  coord_equal() +
  scale_color_brewer(type = "qual", palette = "Dark2") +
  geom_text(fontface = "bold") +
  labs(color="River", x="DCA 1", y="DCA 2", 
       title = "Detrended correspondence analysis", subtitle = "by sample_type") +
  theme(legend.position = "none")
```

###3 d 
```{r}
library(vegan3d)
vegan3d::ordirgl(object = spp_rct_dca, display = "sites", type = "p")
```


