---
title: "Riparian Invertebrates: Accumulation curves parameterised"
output: html_notebook
---

##Packages
```{r, message=FALSE, warning=FALSE}
# packages
library(tidyverse)
library(readxl)
library(forcats)
library(vegan)
```

##Import data
```{r}
#effort_data <- read_csv("../data/effort.csv") 
field_data <- read_csv("../data/field_data_selected.csv")
field_data <-  field_data %>% 
  #get rid of event = NA
  filter(!is.na(event))

#indices <- read_csv("../data/indices.csv")
#river_indices <- read_csv("../data/river_indices.csv")
```




Lets convert this to a frequency matrix
```{r}
field_data_freq <- field_data %>% 
  #get rid of event = NA
  filter(!is.na(event)) %>% 
  #calculate frequencies
  group_by(river, event, sample_type, spp_name) %>% 
  summarise(n = length(spp_name)) %>% 
  #abbreviate species names
  mutate(spp_name = vegan::make.cepnames(spp_name)) %>% 
  #spread to a matrix
  spread(key = spp_name, value = n, fill = 0) %>%
  data.frame() %>%
  arrange(river) 

field_data_freq
```


```{r}
#Parameters
#param_sampletype <- c("excavation", "hand_search", "pitfall")
#param_event <- unique(field_data$event)
param_river <- NULL

```

species accumulation curve by sample type
```{r}
param_sampletype <- c("excavation", "hand_search", "pitfall")

#create empty data frame
spec_accum_all <- data_frame(
  sample_type = as.character(), 
  n_sites = as.integer(),
  n_spp = as.double(),
  sd = as.double())

#iterate through parameters
for(sampletype in param_sampletype){

  # #for testing, assign single value to parameter
  # sampletype <- "pitfall"
  
  tempdata <- field_data_freq %>%
    filter(sample_type == sampletype) %>% 
    select(-c(river, event, sample_type)) %>% 
    specaccum() 
  
  tempdata <- data.frame(
      sample_type = sampletype,
      n_sites = tempdata$sites,
      n_spp = tempdata$richness,
      sd = tempdata$sd) 
  
  spec_accum_all <- rbind(spec_accum_all, tempdata)
  
  rm(tempdata)
}


ggplot(spec_accum_all, aes(x = n_sites, y = n_spp, colour = sample_type))+
  #geom_path(aes(y = n_spp-sd), alpha = 0.5) +
  #geom_path(aes(y = n_spp+sd), alpha = 0.5) +
  #geom_linerange(aes(ymin = n_spp-sd, ymax = n_spp+sd)) +
  geom_path(size = 1) +
  labs(title = "Species accumulation curve by sample type",
       subtitle = paste0(param_sampletype, collapse = ", "),
       caption = "", 
       #colour = "", 
       #fill = "", 
       #x = "", 
       #y = "",
       NULL) +
  NULL

```




```{r}
param_river <- c("Severn", "Wye", "Towy")  #unique(field_data$river)
  

#create empty data frame
spec_accum_all <- data_frame(
  river = as.character(), 
  n_sites = as.integer(),
  n_spp = as.double(),
  sd = as.double())

#iterate through parameters
for(river_i in param_river){

  tempdata <- field_data_freq %>%
    filter(river == river_i) %>% 
    select(-c(river, event, sample_type)) %>% 
    specaccum() 
  
  tempdata <- data.frame(
      river = river_i,
      n_sites = tempdata$sites,
      n_spp = tempdata$richness,
      sd = tempdata$sd) 
  
  spec_accum_all <- rbind(spec_accum_all, tempdata)
  
  rm(tempdata)
}


ggplot(spec_accum_all, aes(x = n_sites, y = n_spp, colour = river))+
  #geom_path(aes(y = n_spp-sd), alpha = 0.5) +
  #geom_path(aes(y = n_spp+sd), alpha = 0.5) +
  #geom_linerange(aes(ymin = n_spp-sd, ymax = n_spp+sd)) +
  geom_path(size = 1) +
  labs(title = "Species accumulation curve by river",
       subtitle = paste0(param_river, collapse = ", "), #"all rivers",
       caption = "", 
       #colour = "", 
       #fill = "", 
       #x = "", 
       #y = "",
       NULL) +
  #theme(legend.position = "none") +
  NULL
#ggsave("../outputs/specaccum_by_river_SvWyTo.png")
```




species accumulation curve for selected surveys
```{r}
library(vegan)

#list of surveys on selected river
events_severn <- unique(field_data$event[which(field_data$river == "Severn")])
events_severn
unique(field_data_freq$event)
table(events_severn %in% field_data_freq$event)
unique(events_severn)
#species accumulation curve by resampling events
spec_accum_severn <- specaccum(comm = field_data_freq %>% 
                                 filter(event %in% events_severn) %>%
                                 select(-c(river, event, sample_type)))
plot(spec_accum_severn, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Severn")
abline(h = max(spec_accum_severn$richness), col = "black")
abline(h = max(spec_accum_severn$richness)*0.95, col = "red")
abline(h = max(spec_accum_severn$richness)*0.9, col = "red", lty = "dotted")
```


```{r}
#list of surveys on selected river
events_towy <- unique(field_data$event[which(field_data$river == "Towy")])
#species accumulation curve by resampling events
spec_accum_towy <- specaccum(comm = field_data_freq %>% 
                                 filter(event %in% events_towy) %>% 
                                 select(-c(river, event, sample_type)))
plot(spec_accum_towy, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Towy")
abline(h = max(spec_accum_towy$richness), col = "black")
abline(h = max(spec_accum_towy$richness)*0.95, col = "red")
abline(h = max(spec_accum_towy$richness)*0.9, col = "red", lty = "dotted")
```


```{r}
#list of surveys on selected river
events_wye <- unique(field_data$event[which(field_data$river == "Wye")])
#species accumulation curve by resampling events
spec_accum_wye <- specaccum(comm = field_data_freq %>% 
                                 filter(event %in% events_wye) %>% 
                                 select(-c(river, event, sample_type)))
plot(spec_accum_wye, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: Wye")
abline(h = max(spec_accum_wye$richness), col = "black")
abline(h = max(spec_accum_wye$richness)*0.95, col = "red")
abline(h = max(spec_accum_wye$richness)*0.9, col = "red", lty = "dotted")
```


```{r}
#param_method <-
events_selected <- c("Wye 2014 5", "Beamish 2013", "Dove 2013", "Lugg 2014 2", "Monnow 2015", "River Frome Moreton 2017", "River Frome Rushton 2017", "River Frome Woodsford 2017", "Till 2013", "Woolder Water 2013", "Wye 2014", "Lugg 2014 1", "Lugg 2014 3", "Lugg 2014 4", "Wye 2014 4", "Wye 2014 6", "Wye 2014 7", "Wye 2014 3")

#species accumulation curve by resampling events
spec_accum_selected <- specaccum(comm = field_data_freq %>% 
                                 filter(event %in% events_selected) %>% 
                                 select(-c(river, event, sample_type)))
plot(spec_accum_selected, ci.type = "polygon", ci.col = "grey", ci.lty = "dotted", 
     main = "Species accumulation curve: selected")
abline(h = max(spec_accum_selected$richness), col = "black")
abline(h = max(spec_accum_selected$richness)*0.95, col = "red")
abline(h = max(spec_accum_selected$richness)*0.9, col = "red", lty = "dotted")


#I want to do as above, but split it by: 
#- author 
#- method: handsearch and pitfall
#- group pitfall by number of pitfalls and plot; group handsearch similarly
#- replicate - i.e. combine date and method 
```




```{r}
plot(spec_accum_wye, col = "green")
plot(spec_accum_severn, add = T, col = "blue")
plot(spec_accum_towy, add = T, col = "orange")

# op <- par()
# par(mfrow = c(1,3))
# plot(spec_accum)
# plot(spec_accum_severn, col = "red", add = T)
#par(op)
```

Try plotting specaccum object in ggplot
```{r}
graphdata <- data.frame(n_sites = spec_accum$sites,
                        n_spp = spec_accum$richness,
                        sd = spec_accum$sd)

ggplot(graphdata, aes(x = n_sites, y = n_spp))+
  geom_point() +
  geom_errorbar(aes(ymin = n_spp-sd, ymax = n_spp+sd))

ggplot(graphdata, aes(x = n_sites, y = n_spp))+
  geom_point() +
  geom_path(aes(y = n_spp-sd)) +
  geom_path(aes(y = n_spp+sd))

plot(spec_accum)
```


#Have a look at inval

* similarity index (jaccard)
* cluster
* labdsv::indval

* simper

* permanova (in vegan see adonis) - permutes anova over multivariate ... aaaagh
  - see which communities are different from one another
    - expl vars e.g. substrate, river, effort; resid variance of site factor
    - can we use pres abs

 
```{r}
labdsv::indval(select(field_data_freq, -c(river, event, sample_type)))
```

* instead of doing species accumulation, do similarity permutation curve



### sample_type
* better to treat the three methods as separate, because wot they catch is different anyway
  - check that with Jon
  - then can do permanova on e.g. pitfalls
  - maybe aggregate river types, have a look at that paper
  - 
  











