---
title: "ripInverts - discussion paper content"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

#Import and prepare data

The following script imports data, cleans it and prepares a number of datasets.  It also carries out a few analyses: 
```{r}
source("../scripts/ripInverts-paper-code_2.R", chdir = T)
```
These are described in more detail below.  

##Data sources

`field_data_selected_Oct2019.csv` contains all individual invertebrate records.  
```{r}
head(obs_all)
skimr::skim_to_wide(obs_all) %>% select(type:sd)
```

```{r}
obs_all %>% select(event, event_code) %>% unique() %>% arrange(event)
```




`speciesmatch_further_correctionsAug2019.csv` is a lookup table of species names: 

```{r}
head(spp_matched_corrected)
skimr::skim_to_wide(spp_matched_corrected)
```

`Riparian Beetle Assemblages July2019 ck added.xlsx` sheet 3 `Effort` is JW calculation of effort scores for each event

```{r}
head(effort_data)
skimr::skim_to_wide(effort_data)
```




```{r}
library(tidyverse)
library(cowplot)
library(iNEXT)
library(vegan)
```

```{r}
#generic graph settings

theme_doc <-  theme(axis.title = element_text(size = 10), 
                    text=element_text(size=10), 
                    plot.title = element_text(size = 11),  
                    legend.key.width = unit(1.2,"cm")) 
```




A total of 


```{r}
#calculate total number of records
nrow(obs_all)
```


records were collected as part of a programme that could be collated into individual  survey events.  A total of 

```{r}
#calculate authors



```

authors generated 

```{r}
#calculate number of events

obs_all %>% filter(is.na(event))

events_unique <- sort(unique(obs_all$event)[which(is.na(unique(obs_all$event))==FALSE)]); events_unique 
length(events_unique)
```


events that occurred between 

```{r}
event_lookup %>% group_by(evt_year) %>% summarise(eventsperyear = n()) 
```




records collected between
```{r}
#calculate date range
table(obs_all$year)

#obs_all %>% filter(year == "1987") %>% count(event, river, location, sample_type)
#obs_all %>% filter(river == "Tees") %>% count(event, river)
#obs_all %>% filter(river == "Wye") %>% count(event, river)
```

, each consisting of one or more location along any given river (e.g. the River Tees consisted of two sampling events in 2008; the River Wye has seven sampling events from 2014) (Table S1). 








## richness and scores

```{r}
#richness by year

#summarise species richness per event
graphdata <- obs_all %>% 
  select(spp_name, event_code, evt_year) %>% 
  unique() %>% 
  group_by(event_code, evt_year) %>% 
  summarise(richness = n()) %>% 
  filter(evt_year != "NA") %>% 
  mutate(evt_year = as.numeric(evt_year))


ggplot(graphdata, aes(x = evt_year, y = richness, group = as.factor(evt_year))) +
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  scale_x_continuous(breaks = seq(1985,2020, by = 1), 
                     labels = c("1985", "", "", "", "", 
                                "1990", "", "", "", "", 
                                "1995", "", "", "", "", 
                                "2000", "", "", "", "", 
                                "2005", "", "", "", "", 
                                "2010", "", "", "", "", 
                                "2015", "", "", "", "", "2020")) +
  scale_y_continuous(breaks = seq(0, 250, by = 25))+
  labs(title = "Species richness in surveys over time",
       x = "Year", y = "Number of species") +
  theme_bw() +
  theme_doc +
  theme(panel.grid.minor.x = element_blank())

#or cropped by leaving out the 1987 records

ggplot(graphdata, aes(x = evt_year, y = richness, group = as.factor(evt_year))) +
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  # geom_violin() +
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  scale_x_continuous(breaks = seq(1997,2020, by = 1), limits = c(1997, 2020), 
                     labels = c("", "", "", "2000", "", "", "", "", 
                                "2005", "", "", "", "", "2010", "", "", "", "", 
                                "2015", "", "", "", "", "2020")) +
  scale_y_continuous(breaks = seq(0, 250, by = 25))+
  labs(title = "Species richness in surveys over time", 
       x = "Year", y = "Number of species") +
  theme_bw() +
  theme_doc +
  theme(panel.grid.minor.x = element_blank())
```

```{r}
#richness by year WETLAND ONLY

#summarise species richness per event
graphdata <- obs_all %>% 
  select(spp_name, event_code, evt_year) %>% 
  left_join(select(effort_data, event_code, effort)) %>% 
  left_join(select(spp_lookup_fam, species, 'Wetland species'), 
            by = c("spp_name" = "species")) %>% 
  rename(wetland_spp = 'Wetland species') %>% 
  filter(wetland_spp == 1) %>% 
  unique() %>% 
  group_by(event_code, evt_year) %>% 
  summarise(richness = n()) %>% 
  filter(evt_year != "NA") %>% 
  mutate(evt_year = as.numeric(evt_year))




ggplot(graphdata, aes(x = evt_year, y = richness, group = as.factor(evt_year))) +
  # geom_boxplot(varwidth = T, colour = "darkgrey") +
  # geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  geom_point(size = 2, colour = "darkgrey") +
  scale_x_continuous(breaks = seq(1997,2020, by = 1), limits = c(1997, 2020), 
                     labels = c("", "", "", "2000", "", "", "", "", 
                                "2005", "", "", "", "", "2010", "", "", "", "", 
                                "2015", "", "", "", "", "2020")) +
  scale_y_continuous(breaks = seq(0, 250, by = 25))+
  labs(title = "Wetland species richness in surveys over time", 
       x = "Year", y = "Wetland species richness") +
  #theme_bw() +
  theme_doc +
  # theme(panel.grid.minor.x = element_blank()) +
  NULL

```



### Effort

```{r}
#effort summarised by JW
effort <- effort_data %>% select(event_code, hand_search = handsearch, pitfall, effort_score = effort); effort

plot_effort <- ggplot(effort, aes(x = effort_score)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(breaks = 0:max(effort$effort_score)) +
  scale_y_continuous(expand = expand_scale(add =c(0,0.5))) +
  labs(x = "Sampling effort", y = "Number of surveys", title = "A") +
  theme_bw()+  theme_doc
plot_effort
```


```{r}
#effort derived from observations data
effort_ck <- obs_all_mat %>% 
  select(event_code, sample_type) %>% 
  group_by(event_code) %>% 
  count(sample_type) %>% 
  spread(key = sample_type, value = n, fill = 0) %>% 
  mutate(effort_score = as.integer(hand_search + pitfall)) %>% 
  filter(event_code != "unknown"); effort %>% arrange(-effort_score)
```
These effort scores are inconsistent with JW's scores.  It is likely that CK has defined 'replicates' in `obs_all_mat` is not an accurate reflection of the way JW grouped sampling instances (e.g. one session of hand searching).  It is not going to be possible to do this accurately without re-grouping all the records from scratch.  

```{r}
wetland_effort <- 
  obs_all %>% select(spp_name, event_code) %>% 
  left_join(select(effort_data, event_code, effort)) %>% 
  left_join(select(spp_lookup_fam, species, 'Wetland species'), 
            by = c("spp_name" = "species")) %>% 
  rename(wetland_spp = 'Wetland species') %>% 
  unique() %>% 
  group_by(event_code, effort) %>% 
  count(wetland_spp) %>% 
  filter(wetland_spp == 1) %>% 
  filter(event_code != "unknown")

wetland_effort %>% arrange(-n)

plot_wetland_effort <- ggplot(wetland_effort, 
                              aes(x = as.factor(effort), y = n), 
                              fill = as.factor(effort))+
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  #geom_violin(scale = "width")+
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  labs(x = "Sampling effort", y = "Number of wetland species", title = "B") +
  theme_bw() +
  theme_doc
  
plot_wetland_effort
```

```{r}
plot_wetland_effort_regr <- ggplot(wetland_effort, 
                              aes(x = effort, y = n), 
                              fill = as.factor(effort))+
  geom_boxplot(varwidth = T, colour = "darkgrey", aes(group = effort)) +
  geom_smooth(method = "lm")+
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  scale_x_continuous(breaks = 2:13) +
  labs(x = "Sampling effort", y = "Number of wetland species", title = "B") +
  theme_bw() + theme_doc; plot_wetland_effort_regr

lm.eff <- lm(effort ~ n, wetland_effort)
broom::tidy(lm.eff)
summary(lm.eff)
lm.eff
```


```{r fig.height=4, fig.width=8}
plot_grid(plot_effort, plot_wetland_effort)
```


##sample type: does it matter
```{r}
adon1
```


## pitfall v handsearch

```{r }

library(ggpubr)
graphdata <-  obs_all_freq_types %>% 
  filter(wetland_spp %in% c(TRUE, FALSE)) %>%
  count(wetland_spp, sample_types) %>%
  group_by(wetland_spp) %>% 
  mutate(prop = n/sum(n))

plotmargin <- unit(c(8, 4, 4, 4), units = "mm")

plot1 <- ggplot(graphdata, aes(x = sample_types, y = n)) +
  geom_bar(aes(alpha = wetland_spp), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey() +
  scale_alpha_discrete(range = c(0.5, 0.9), labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.x = element_blank(),
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c(h = "handsearch", hp = "both", p = "pitfall")) +
  labs(#title = "Does capture method matter?", 
       x = "", 
       y = "number of species",
       alpha = "species type")
plot1

plot2 <- ggplot(graphdata, aes(x = wetland_spp, y = prop)) +
  geom_bar(aes(fill = sample_types), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey(start = 0.8, end = 0.3, 
                  labels = c(h = "handsearch", hp = "both", p = "pitfall"))+
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.y = element_blank(), 
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  labs(#title = "Does capture method matter?", 
       x = "species type", 
       y = "proportion of species",
       fill = "sample type")
plot2

title <- ggdraw() + draw_label("Is pitfall trapping alone sufficient?", fontface='bold')
caption <- ggdraw() + draw_text("There is no relationship between the type of species captured (whether wetland or not wetland species) and the method of capture.  Fig x.A shows the total number of each type of species by capture method and fig x.B shows the proportion of species captured with each capture method, split by wetland and non-wetland species")


plot_grid(title, plot1, plot2 + coord_flip(),
          nrow = 3, labels = c("", "A", "B"), 
          rel_heights = c(0.6, 1.5, 1), 
          vjust = -0 )  

```

what proportion of wetland species are caught by handsearch only?

keep top one only.  

add boxplot back in


```{r}
sample_type_count <- obs_all %>% 
  select(river, event_code, sample_type, replicate, year) %>%
  group_by(event_code, replicate, sample_type) %>% 
  count(sample_type) %>% 
  ungroup() %>% 
  select(event_code, sample_type) %>% 
  count(event_code, sample_type) %>% 
  spread(sample_type, n)

sample_type_count

paste("there are", length(unique(obs_all$replicate)), "replicates in total")

sample_type_count %>% select(hand_search, pitfall) %>% colSums(na.rm = T)

```



```{r}
graphdata <-  obs_all_freq_types %>% 
  count(wetland_spp, sample_types) %>%
  group_by(wetland_spp) %>% 
  mutate(prop = n/sum(n))
graphdata

sum(graphdata$n)

plotmargin <- unit(c(8, 4, 4, 4), units = "mm")


ggplot(graphdata, aes(x = wetland_spp, y = n)) +
  geom_bar(aes(fill = sample_types), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(h = "handsearch only", hp = "handsearch & \npitfall", 
                             p = "pitfall only"))+
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.x = element_blank(), 
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  labs(title = "Pitfalls, Handsearch or Both?", 
       x = "species type", 
       y = "number of species captured",
       fill = "sample type") +
  theme_doc


```



```{r}
chidata <- obs_all_freq_types %>%
  group_by(wetland_spp) %>%
  count(sample_types) %>%
  filter(wetland_spp %in% c(TRUE, FALSE)) %>%
  spread(key = wetland_spp, value = n) %>%
  mutate(sample_types = recode_factor(sample_types, 
                                      h = "handsearch", hp = "both", p = "pitfall")) %>% 
  column_to_rownames("sample_types") 

chidata  
#note that the species numbers here exclude any that are not classified as either wetland or not-wetland.  Perhaps ought to classify all NAs as not wetland.  

chisqresult <- chisq.test(chidata)

chisqresult

chisqresult$observed
sum(chisqresult$observed)
totals <- chisqresult$observed %>% data.frame() %>%  colSums(); totals

chisqresult$observed[,1] / totals[1]
chisqresult$observed[,2] / totals[2]
```

```{r}
data.frame(h = c(65, 95), hpp = c(160, 205))
chisq.test(data.frame(h = c(65, 95), hpp = c(160, 205)))
```
### compare JW to CK wetland

```{r}
wetland_effort
jw_spp_event_data <- jw_spp_event_data %>%   mutate(event_code = Event_code_v3) %>% 
  filter(is.na(`New Name`) == FALSE)
jw_spp_event_data

left_join(wetland_effort, jw_spp_event_data) %>% 
  ungroup() %>% 
  mutate(event_code, 
         n_wet_ck = n, 
         n_wet_jw = `Total Number of Wetland Species`, 
         diff = n_wet_ck - n_wet_jw) %>% 
  select(event_code, n_wet_ck, n_wet_jw, diff) %>%
  arrange(-abs(diff)) %>% 
  print() %>% write_csv("../data/wetland_compare.csv")
```



###spp lenght

```{r}
ggplot(obs_all_freq_types, aes(x = length.min)) +
  geom_histogram(aes(fill = sample_types), colour = "black",
                 stat = "bin", binwidth = 0.5) +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(h = "handsearch", hp = "both", p = "pitfall"))+
  theme_minimal() +
  theme(legend.position = c(0.9, 0.8))+
  labs(title = "Distribution of species \nby capture type and minimum length", 
       y = "number of species", x = "minimum length (mm)", fill = "capture type") +
  theme_doc

```


###by genus

```{r}
genus_levels <-  samptype_genus$genus[order(-samptype_genus$n_spp)]
min_n_spp <- 4

graphdata <- samptype_genus %>%
  ungroup() %>% 
  filter(n_spp > min_n_spp) %>% 
  arrange(-n_spp) %>%
  mutate(genus = factor(genus, levels = rev(genus))) %>% 
  select(-n_spp) %>% 
  gather(key = sample_types, 
         value = n_spp, handsearch, pitfall, handsearch_pitfall) %>% 
  droplevels()

n_genus <- length(unique(graphdata$genus))

ggplot(graphdata, aes(y = n_spp, x = genus, fill = sample_types)) +
  geom_col(colour = "black", width = 0.7) +
  coord_flip() +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(handsearch = "handsearch", handsearch_pitfall = "both",
                             pitfall = "pitfall")) +
  theme_minimal() +
  theme(legend.position = c(0.8, 0.5)) +
  labs(title = "Capture methods by genus", 
       caption = paste0("Method of capture for the ", n_genus, 
                        " most fequently captured genera."),
       y = "Number of species", x = "", fill = "sample type") +
  theme_doc
```

### by family

```{r}
samptype_fam

family_levels <-  samptype_fam$Family[order(-samptype_fam$n_spp)]
min_n_spp <- 4

graphdata <- samptype_fam %>%
  ungroup() %>% 
  filter(n_spp > min_n_spp) %>% 
  arrange(-n_spp) %>%
  mutate(Family = factor(Family, levels = rev(Family))) %>% 
  select(-n_spp) %>% 
  gather(key = sample_types, 
         value = n_spp, handsearch, pitfall, handsearch_pitfall) %>% 
  droplevels()

n_family <- length(unique(graphdata$Family))

ggplot(graphdata, aes(y = n_spp, x = Family, fill = sample_types)) +
  geom_col(colour = "black", width = 0.7) +
  coord_flip() +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(handsearch = "handsearch", handsearch_pitfall = "both",
                             pitfall = "pitfall")) +
  theme_minimal() +
  theme(legend.position = c(0.8, 0.5)) +
  labs(title = "Capture methods by family", 
       caption = paste0("Method of capture for the ", n_family, 
                        " most fequently captured families."),
       y = "Number of species", x = "", fill = "sample type") +
  theme_doc

```


## species accumulation

```{r}
graphdata <- spec_accum_byriver_evt 
legenddata <- graphdata %>% group_by(river) %>% summarise(max_sites = max(n_sites)) 
caption <- 
"Species accumulation curve using the 'exact' method of the specaccum function in 
R package vegan (Oksanen et al. 2019).  After calculating species richness of all 
permutations of the surveys, this graph shows the average number of species found 
for a given number of surveys, with standard error bars. The error reduces to zero 
at the total number of surveys as this equals the total number of species found."

plot_accum <- ggplot(graphdata, aes(x = n_sites, y = n_spp, colour = river))+
  geom_line(size = 1) +
  geom_errorbar(aes(ymin=n_spp-(sd/sqrt(n_sites)), ymax=n_spp+(sd/sqrt(n_sites))),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.1)) +  
  scale_colour_grey(start = 0.4, end = 0.8, 
                        labels = paste0(legenddata$river, " (", 
                                        legenddata$max_sites, ")"))+
  labs(title = "Observed species accumulation", 
       x = "number of surveys", y = "number of species") +
  theme_bw() +
  theme(legend.position = c(0.7, 0.2), legend.title = element_blank()) +
  theme_doc +
  NULL; plot_accum

print(caption)
```

```{r}
spec_accum_byriver_evt %>% filter(sd == 0)
```

THIS SEEMS TO BE BROKEN.  I don't know where and how I created `eventdataList`

```{r}
library(iNEXT)
data(ciliates)

data.frame(ciliates$EtoshaPan)
```


```{r}
library(iNEXT)
library(tidyverse)
eventdataList <- list()
  
eventdataList <- list(
  "Frome" = obs_rct_mat_evt %>% filter(river == "Frome") %>% select(-river) %>%
  column_to_rownames(var = "event_code") %>%  as.matrix() %>% t(),
  "Lugg" = obs_rct_mat_evt %>% filter(river == "Lugg") %>% select(-river) %>%
  column_to_rownames(var = "event_code") %>%  as.matrix() %>% t(),
  "Wye" = obs_rct_mat_evt %>% filter(river == "Wye") %>% select(-river) %>%
  column_to_rownames(var = "event_code") %>%  as.matrix() %>% t()
  )
  
str(eventdataList)
```





```{r fig.height=5.5, fig.width=4}

eventdata_iNext <- iNEXT(eventdataList, datatype = "incidence_raw", q = 0, endpoint = 40, knots = 40)

asympdata <- eventdata_iNext$AsyEst %>% filter(Diversity == "Species richness")

inext_output <- bind_rows(eventdata_iNext$iNextEst, .id = "Site") %>%
  mutate(method2 = recode_factor(method, observed = "interpolated")) %>%
  full_join(asympdata)

inext_output

effort_95 <- inext_output %>% filter(SC >= 0.95) %>% group_by(Site) %>% top_n(n = 1,wt = -qD)
effort_95
```

Notes on iNEXT output: 

t = sample size for which diversity estimates of order q are computde
method = interpolated, observed, or extrapolated, depending on whether the size t is less than, equal to, or greater than the reference sample size.
order = the diversity order of q	
qD = the estimated diversity of order q for a sample of size t.
SC = the estimated sample coverage for a sample of size t.
qD.LCL, qD.UCL = the bootstrap lower and upper confidence limits for the diversity of order q at the specified level in the setting (with a default value of 0.95).
SC.LCL, SC.UCL = the bootstrap lower and upper confidence limits

```{r fig.height=8, fig.width=5}
par.orig <- par()
par(mfrow = c(3,1))
plot(eventdata_iNext,type = 1)
plot(eventdata_iNext,type = 2)
plot(eventdata_iNext, type = 3)
```


```{r}
caption <-
  "Extrapolation of species accumulation using R function iNEXT (Hsieh et al., 2019).
This extrapolates a curve of expected species richness for each additional survey
beyond the maximum number of actual surveys, guided by an estimated asymptote (Chao
et al., 2014).  Hill number q = 0 was used as a measure of diversity."

plot_asympt <- ggplot(inext_output, aes(x = t)) +
  #plot extrapolated and interpolated data
  geom_line(aes(y = qD, linetype = method2), size = 1) +
  geom_ribbon(aes(ymin=qD.LCL, ymax=qD.UCL), alpha=0.2) +
  #plot guides for 95% of richness
  geom_segment(data = effort_95, aes(x = 0, xend = t, y = qD, yend = qD),
               alpha = 0.6, size = 0.1, linetype = 18)+
  geom_segment(data = effort_95, aes(x = t, xend = t, y = 0, yend = qD),
               alpha = 0.6, size = 0.1, linetype = 18)+
  #formatting
  scale_color_grey() +
  scale_fill_grey() +
  scale_linetype_manual(values = c(1,3)) +
  facet_wrap(~Site,ncol = 1) +
  labs(title = "Extrapolated species richness",
       y = "number of species", x = "number of surveys") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  coord_cartesian(xlim = c(1, 40)) +
  scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0))+
  theme_doc +
  NULL; plot_asympt

print(caption)


```


```{r}
#prepare tables

inext_output %>% filter(SC >= 0.95) %>% group_by(Site) %>% top_n(n = 5,wt = -qD)

#eventdata_iNext$iNextEst$Frome
eventdata_iNext$AsyEst


```
```{r}
#find the nearest number of samples where Sample Coverage is at 0.95 and same for SC.LCL and UCL 
inext_output %>% group_by(Site) %>% 
  summarise(t.95 = t[which(SC == SC[which.min(abs(SC - 0.95))])],
            t.95.LCL = t[which(SC.UCL == SC.UCL[which.min(abs(SC.UCL - 0.95))])],
            t.95.UCL = t[which(SC.LCL == SC.LCL[which.min(abs(SC.LCL - 0.95))])])
  
  
  
```


```{r}
write.csv(as.data.frame(eventdata_iNext$AsyEst), "../data/asympt_output.csv")
write.csv(as.data.frame(inext_output), "../data/inext_output.csv")
```



```{r}
citation("iNEXT")
print(citation("iNEXT"), bibtex=TRUE)
```



```{r}
library(cowplot)

plot_grid(plot_accum, plot_asympt)

```

## Look at genus again


```{r}
genusdata <- obs_all_freq_types %>% 
  separate(col = spp_name, into = c("genus", NA), remove = F) %>% 
  group_by(genus) %>% 
  mutate(pitfall = -pitfall) %>% 
  arrange(spp_name) %>% 
  gather(key = sample_type, value = n_samples, hand_search, pitfall)
```

```{r fig.height=3, fig.width=6}
selected_genera <- c("Bledius")#, "Bembidion")
graphdata <- filter(genusdata, genus %in% selected_genera)
minlabel <- round(min(graphdata$n_samples)/5)*5
maxlabel <- round(max(graphdata$n_samples)/5)*5
breaks <- seq(minlabel, maxlabel, 5)

ggplot(graphdata, aes(x = spp_name, y = n_samples, alpha = sample_type))+
  geom_col(width = 0.6) +
  scale_alpha_manual(values = c("pitfall" = 0.55, "hand_search" = 0.85)) +
  scale_y_continuous(breaks = breaks, labels = abs(breaks), minor_breaks = NULL) +
  annotate(geom = "text", label = c("<- pitfall  handsearch ->"), family = "mono",
           x = length(unique(graphdata$spp_name))+0.5, y = 1, size = 3) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme_minimal() +  theme(legend.position = "none", axis.title.y = element_blank()) +  coord_flip()
```



```{r}
selected_genera <- c("Bembidion")
graphdata <- filter(genusdata, genus %in% selected_genera)
minlabel <- round(min(graphdata$n_samples)/50)*50
maxlabel <- round(max(graphdata$n_samples)/50)*50
breaks <- seq(minlabel, maxlabel, 50)

ggplot(graphdata, aes(x = spp_name, y = n_samples, alpha = sample_type))+
  geom_col(width = 0.7) +
  scale_alpha_manual(values = c("pitfall" = 0.55, "hand_search" = 0.85)) +
  scale_y_continuous(breaks = breaks, labels = abs(breaks), minor_breaks = NULL) +
  annotate(geom = "text", label = c("<- pitfall  handsearch ->"), family = "mono",
           x = length(unique(graphdata$spp_name))+0.5, y = 10, size = 3) +
  geom_hline(yintercept = 0, alpha = 0.5) +
  theme_minimal() +  theme(legend.position = "none", axis.title.y = element_blank()) +  coord_flip()
```


```{r}
selected_genera <- c("Bembidion", "Bledius", "Hydrosmecta")
graphdata <- filter(genusdata, genus %in% selected_genera)
minlabel <- round(min(graphdata$n_samples)/50)*50
maxlabel <- round(max(graphdata$n_samples)/50)*50
breaks <- seq(minlabel, maxlabel, 50)

ggplot(graphdata, aes(x = spp_name, y = n_samples, alpha = sample_type, fill = genus))+
  geom_col(width = 0.7) +
  scale_alpha_manual(values = c("pitfall" = 0.55, "hand_search" = 0.85)) +
  scale_y_continuous(breaks = breaks, labels = abs(breaks), minor_breaks = NULL) +
  annotate(geom = "text", label = c("<- pitfall  handsearch ->"), family = "mono",
           x = length(unique(graphdata$spp_name))+0.5, y = 10, size = 3) +
  geom_hline(yintercept = 0, alpha = 0.5) +
#  facet_wrap(~genus, scales = "free_x")+
  theme_minimal() +  
  theme(legend.position = "none", axis.title.y = element_blank(), axis.text.y = element_text(size = 7)) +  coord_flip() +
  NULL
```


```{r}
selected_genera <- c("Bembidion", "Bledius", "Hydrosmecta", "Stenus")
graphdata <- filter(genusdata, genus %in% selected_genera)
graphdata$genus <- as.factor(graphdata$genus)
graphdata %>% arrange(n_samples)
minlabel <- round(min(graphdata$n_samples)/50)*50
maxlabel <- round(max(graphdata$n_samples)/50)*50
breaks <- seq(minlabel, maxlabel, 50)

ggplot(graphdata, aes(x = spp_name, y = n_samples, group = spp_name))+#, alpha = sample_type, fill = genus))+
  geom_col(position = position_dodge(width = 10, preserve = "single"), width = 1,
           aes(alpha = sample_type, fill = genus)) +
  scale_alpha_manual(values = c("pitfall" = 0.55, "hand_search" = 0.85)) +
  scale_y_continuous(breaks = breaks, labels = abs(breaks), minor_breaks = NULL) +
  scale_x_discrete()+#labels = element_blank()) +
  annotate(geom = "text", label = c("<- pitfall  handsearch ->"), family = "mono",
           x = 0.5, y = 10, size = 3) +

  #          x = length(unique(graphdata$spp_name))+0.5, y = 10, size = 3) +
  geom_hline(yintercept = 0, alpha = 0.5) +
#  facet_wrap(~genus, scales = "free_x")+
  theme_minimal() +  theme(#legend.position = "none", 
                           axis.title.y = element_blank(),
                           axis.text.y = element_text(size = 7)) +  
  coord_flip() +
  NULL

```


### Species list

```{r}
species_list <- unique(obs_all$spp_name)

```


### by Family


```{r}
spp_lookup_fam %>% 
  filter(species %in% species_list) %>% 
  group_by(`Wetland species`) %>% count(Family) %>% spread(key = `Wetland species`, value = n, fill = 0) %>% 
  select(Family, wetland = `1`, not_wetland =`0`) %>% mutate(total = not_wetland + wetland) %>% arrange(-total, -wetland) %>% print() %>%
  write_csv("../data/family_wetland.csv")
```




