---
title: "ripInverts - discussion paper content"
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r}
source("../scripts/ripInverts-paper-code_1.R", chdir = T)
```

```{r}
library(tidyverse)
library(cowplot)
library(iNEXT)
library(vegan)
```

```{r}
#generic graph settings

theme_doc <-  theme(axis.title = element_text(size = 10), 
                    text=element_text(size=10), 
                    plot.title = element_text(size = 11),  
                    legend.key.width = unit(1.2,"cm")) 
```




A total of 


```{r}
#calculate total number of records
nrow(obs_all)
```


records were collected as part of a programme that could be collated into individual  survey events.  A total of 

```{r}
#calculate authors



```

authors generated 

```{r}
#calculate number of events

obs_all %>% filter(is.na(event))

unique(obs_all$event)[which(is.na(unique(obs_all$event))==FALSE)]
length(unique(obs_all$event)[which(is.na(unique(obs_all$event))==FALSE)])
```


events that occurred between 

```{r}
event_lookup %>% group_by(evt_year) %>% summarise(eventsperyear = n()) 
```




records collected between
```{r}
#calculate date range
table(obs_all$year)

#obs_all %>% filter(year == "1987") %>% count(event, river, location, sample_type)
#obs_all %>% filter(river == "Tees") %>% count(event, river)
#obs_all %>% filter(river == "Wye") %>% count(event, river)
```

, each consisting of one or more location along any given river (e.g. the River Tees consisted of two sampling events in 2008; the River Wye has seven sampling events from 2014) (Table S1). 








##grid references
```{r}
obs_all %>% count(gridref) %>% arrange(-n)
```

```{r}
obs_all %>% count(event_code, coord_precision)
```


```{r}
ggplot(obs_all, aes(easting, northing)) +
  geom_point(aes(colour = event)) + coord_equal() + theme(legend.position = "none")
```


## richness and scores

```{r}
#richness by year

#summarise species richness per event
graphdata <- obs_all %>% 
  select(spp_name, event_code, evt_year) %>% 
  unique() %>% 
  group_by(event_code, evt_year) %>% 
  summarise(richness = n()) %>% 
  filter(evt_year != "NA") %>% 
  mutate(evt_year = as.numeric(evt_year))


ggplot(graphdata, aes(x = evt_year, y = richness, group = as.factor(evt_year))) +
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  scale_x_continuous(breaks = seq(1985,2020, by = 5)) +
  scale_y_continuous(breaks = seq(0, 250, by = 25))+
  labs(title = "Species richness in surveys over time",
       x = "Year", y = "Number of species") +
  theme_bw() +
  theme_doc 

```

### Effort

```{r}
effort <- obs_all_mat %>% 
  select(event_code, sample_type) %>% 
  group_by(event_code) %>% 
  count(sample_type) %>% 
  spread(key = sample_type, value = n, fill = 0) %>% 
  mutate(effort_score = as.integer(hand_search + pitfall)) %>% 
  filter(event_code != "unknown"); effort

effort %>% arrange(-effort_score)

plot_effort <- ggplot(effort, aes(x = effort_score)) +
  geom_histogram(binwidth = 0.5) +
  scale_x_continuous(breaks = 0:max(effort$effort_score)) +
  scale_y_continuous(breaks = seq(0,30, by = 5), expand = c(0,0)) +
  labs(x = "Sampling effort", y = "Number of surveys", title = "A") +
  theme_bw()+
  theme_doc

plot_effort
```


```{r}
wetland_effort <- 
  obs_all %>% select(spp_name, event_code) %>% 
  left_join(select(effort, event_code, effort_score)) %>% 
  left_join(select(spp_lookup, species, 'Wetland species'), 
            by = c("spp_name" = "species")) %>% 
  rename(wetland_spp = 'Wetland species') %>% 
  unique() %>% 
  group_by(event_code, effort_score) %>% 
  count(wetland_spp) %>% 
  filter(wetland_spp == 1) %>% 
  filter(event_code != "unknown")

wetland_effort %>% arrange(-n)

plot_wetland_effort <- ggplot(wetland_effort, 
                              aes(x = as.factor(effort_score), y = n), 
                              fill = as.factor(effort_score))+
  geom_boxplot(varwidth = T, colour = "darkgrey") +
  #geom_violin(scale = "width")+
  geom_point(position = position_jitter(width = 0.08, height = 0.3), size = 0.9)+
  labs(x = "Sampling effort", y = "Number of wetland species", title = "B") +
  theme_bw() +
  theme_doc
  
plot_wetland_effort
```

```{r fig.height=4, fig.width=8}
plot_grid(plot_effort, plot_wetland_effort)
```


## pitfall v handsearch

```{r }

library(ggpubr)
graphdata <-  obs_all_freq_types %>% 
  filter(wetland_spp %in% c(TRUE, FALSE)) %>%
  count(wetland_spp, sample_types) %>%
  group_by(wetland_spp) %>% 
  mutate(prop = n/sum(n))

plotmargin <- unit(c(8, 4, 4, 4), units = "mm")

plot1 <- ggplot(graphdata, aes(x = sample_types, y = n)) +
  geom_bar(aes(alpha = wetland_spp), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey() +
  scale_alpha_discrete(range = c(0.5, 0.9), labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.x = element_blank(),
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c(h = "handsearch", hp = "both", p = "pitfall")) +
  labs(#title = "Does capture method matter?", 
       x = "", 
       y = "number of species",
       alpha = "species type")


plot2 <- ggplot(graphdata, aes(x = wetland_spp, y = prop)) +
  geom_bar(aes(fill = sample_types), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey(start = 0.8, end = 0.3, 
                  labels = c(h = "handsearch", hp = "both", p = "pitfall"))+
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.y = element_blank(), 
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  labs(#title = "Does capture method matter?", 
       x = "species type", 
       y = "proportion of species",
       fill = "sample type")

title <- ggdraw() + draw_label("Is pitfall trapping alone sufficient?", fontface='bold')
caption <- ggdraw() + draw_text("There is no relationship between the type of species captured (whether wetland or not wetland species) and the method of capture.  Fig x.A shows the total number of each type of species by capture method and fig x.B shows the proportion of species captured with each capture method, split by wetland and non-wetland species")


plot_grid(title, plot1, plot2 + coord_flip(),
          nrow = 3, labels = c("", "A", "B"), 
          rel_heights = c(0.6, 1.5, 1), 
          vjust = -0 )  

```

what proportion of wetland species are caught by handsearch only?

keep top one only.  

add boxplot back in


```{r}
chitest_wetland_props$method
```

```{r}
sample_type_count <- obs_all %>% 
  select(river, event_code, sample_type, replicate, year) %>%
  group_by(event_code, replicate, sample_type) %>% 
  count(sample_type) %>% 
  ungroup() %>% 
  select(event_code, sample_type) %>% 
  count(event_code, sample_type) %>% 
  spread(sample_type, n)

sample_type_count

paste("there are", length(unique(obs_all$replicate)), "replicates in total")

sample_type_count %>% select(hand_search, pitfall) %>% colSums(na.rm = T)

```



```{r}
graphdata <-  obs_all_freq_types %>% 
  filter(wetland_spp %in% c(TRUE, FALSE)) %>%
  count(wetland_spp, sample_types) %>%
  group_by(wetland_spp) %>% 
  mutate(prop = n/sum(n))
graphdata

plotmargin <- unit(c(8, 4, 4, 4), units = "mm")


ggplot(graphdata, aes(x = wetland_spp, y = n)) +
  geom_bar(aes(fill = sample_types), position = position_stack(), 
           stat = "identity", colour = "black") +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(h = "handsearch only", hp = "handsearch & \npitfall", 
                             p = "pitfall only"))+
  theme_minimal() +
  theme(legend.position = "right", panel.grid.major.x = element_blank(), 
        plot.margin = plotmargin) +
  scale_x_discrete(labels = c("FALSE" = "not wetland", "TRUE" = "wetland")) +
  labs(title = "Pitfalls, Handsearch or Both?", 
       x = "species type", 
       y = "number of species captured",
       fill = "sample type") +
  theme_doc


```



```{r}
chidata <- obs_all_freq_types %>% 
  filter(wetland_spp %in% c(TRUE, FALSE)) %>% 
  mutate(wetland_spp = as.character(wetland_spp)) %>%
  mutate(wetland_spp = recode(wetland_spp, 
                              "FALSE" = "not-wetland", "TRUE" = "wetland")) %>% 
  group_by(wetland_spp) %>% 
  count(sample_types) %>% 
  spread(key = wetland_spp, value = n) %>%
  mutate(sample_types = recode_factor(sample_types, 
                                      h = "handsearch", hp = "both", p = "pitfall")) %>% 
  column_to_rownames("sample_types") 

chidata  


chisqresult <- chisq.test(chidata)

chisqresult

chisqresult$observed
```

```{r}
data.frame(h = c(65, 95), hpp = c(160, 205))
chisq.test(data.frame(h = c(65, 95), hpp = c(160, 205)))
```


```{r}
ggplot(obs_all_freq_types, aes(x = length.min)) +
  geom_histogram(aes(fill = sample_types), colour = "black",
                 stat = "bin", binwidth = 0.5) +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(h = "handsearch", hp = "both", p = "pitfall"))+
  theme_minimal() +
  theme(legend.position = c(0.9, 0.8))+
  labs(title = "Distribution of species \nby capture type and minimum length", 
       y = "number of species", x = "minimum length (mm)", fill = "capture type") +
  theme_doc

```




```{r}
genus_levels <- genus_types$genus[order(-genus_types$n_spp)]
min_n_spp <- 4

graphdata <- genus_types %>%
  ungroup() %>% 
  filter(n_spp > min_n_spp) %>% 
  arrange(-n_spp) %>%
  mutate(genus = factor(genus, levels = rev(genus))) %>% 
  select(-n_spp) %>% 
  gather(key = sample_types, 
         value = n_spp, handsearch, pitfall, handsearch_pitfall) %>% 
  droplevels()

n_genus <- length(unique(graphdata$genus))

ggplot(graphdata, aes(y = n_spp, x = genus, fill = sample_types)) +
  geom_col(colour = "black", width = 0.7) +
  coord_flip() +
  scale_fill_grey(start = 0.8, end = 0.3,
                  labels = c(handsearch = "handsearch", handsearch_pitfall = "both",
                             pitfall = "pitfall")) +
  theme_minimal() +
  theme(legend.position = c(0.8, 0.5)) +
  labs(title = "Capture methods by genus", 
       caption = paste0("Method of capture for the ", n_genus, 
                        " most fequently captured genera."),
       y = "Number of species", x = "", fill = "sample type") +
  theme_doc
```


### species accumulation



```{r}
graphdata <- spec_accum_byriver_evt 
legenddata <- graphdata %>% group_by(river) %>% summarise(max_sites = max(n_sites)) 
caption <- 
"Species accumulation curve using the 'exact' method of the specaccum function in 
R package vegan (Oksanen et al. 2019).  After calculating species richness of all 
permutations of the surveys, this graph shows the average number of species found 
for a given number of surveys, with standard error bars. The error reduces to zero 
at the total number of surveys as this equals the total number of species found."

plot_accum <- ggplot(graphdata, aes(x = n_sites, y = n_spp, colour = river))+
  geom_line(size = 1) +
  geom_errorbar(aes(ymin=n_spp-(sd/sqrt(n_sites)), ymax=n_spp+(sd/sqrt(n_sites))),
                  size=.3,    # Thinner lines
                  width=.2,
                  position=position_dodge(.1)) +  
  scale_colour_grey(start = 0.4, end = 0.8, 
                        labels = paste0(legenddata$river, " (", 
                                        legenddata$max_sites, ")"))+
  labs(title = "Observed species accumulation", 
       x = "number of surveys", y = "number of species") +
  theme_bw() +
  theme(legend.position = c(0.7, 0.2), legend.title = element_blank()) +
  theme_doc +
  NULL; plot_accum

print(caption)
```

```{r}
spec_accum_byriver_evt %>% filter(sd == 0)
```



```{r fig.height=5.5, fig.width=4}

eventdata_iNext <- iNEXT(eventdataList, datatype = "incidence_raw", q = 0, endpoint = 40, knots = 40)

asympdata <- eventdata_iNext$AsyEst %>% filter(Diversity == "Species richness")

graphdata <- bind_rows(eventdata_iNext$iNextEst, .id = "Site") %>% 
  mutate(method2 = recode_factor(method, observed = "interpolated")) %>% 
  full_join(asympdata)

graphdata

effort_95 <- graphdata %>% filter(SC >= 0.95) %>% group_by(Site) %>% top_n(n = 1,wt = -qD)
effort_95

caption <- 
  "Extrapolation of species accumulation using R function iNEXT (Hsieh et al., 2019).  
This extrapolates a curve of expected species richness for each additional survey 
beyond the maximum number of actual surveys, guided by an estimated asymptote (Chao 
et al., 2014).  Hill number q = 0 was used as a measure of diversity."

plot_asympt <- ggplot(graphdata, aes(x = t)) +
  #plot extrapolated and interpolated data
  geom_line(aes(y = qD, linetype = method2), size = 1) +
  geom_ribbon(aes(ymin=qD.LCL, ymax=qD.UCL), alpha=0.2) +
  #plot guides for 95% of richness
  geom_segment(data = effort_95, aes(x = 0, xend = t, y = qD, yend = qD), 
               alpha = 0.6, size = 0.1, linetype = 18)+
  geom_segment(data = effort_95, aes(x = t, xend = t, y = 0, yend = qD), 
               alpha = 0.6, size = 0.1, linetype = 18)+  
  #formatting
  scale_color_grey() +
  scale_fill_grey() +
  scale_linetype_manual(values = c(1,3)) +
  facet_wrap(~Site,ncol = 1) +
  labs(title = "Extrapolated species richness", 
       y = "number of species", x = "number of surveys") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  coord_cartesian(xlim = c(1, 40)) +
  scale_x_continuous(expand = c(0,0)) + scale_y_continuous(expand = c(0,0))+
  theme_doc +
  NULL; plot_asympt

print(caption)


```




```{r}
write.csv(as.data.frame(eventdata_iNext$AsyEst), "../data/asympt_output.csv")
write.csv(as.data.frame(graphdata), "../data/inext_output.csv")
```



```{r}
citation("iNEXT")
print(citation("iNEXT"), bibtex=TRUE)
```



```{r}
library(cowplot)

plot_grid(plot_accum, plot_asympt)

```



