---
title: "Riparian Invertebrates: Data Import 4"
output: html_notebook
---

This script consolidates dataImport_2 and dataImport3, and replaces some of teh data with a new version received from JW in March 2019.  

##Packages
```{r}
# packages
library(tidyverse)
library(readxl)
library(forcats)
```


##Import data 
```{r}
#importing all as text to begin with
biol_records <- read_csv("../data/Riparian Event Data Jan2019 - biological records.csv")

events_code <- read_csv("../data/event_code_lookup_orig.csv") %>% select(-1)
```



```{r}
#summary
n_spp <- length(unique(biol_records$Species_name))
n_rivers <- length(unique(biol_records$`Sample River`))
n_events <-  length(unique(biol_records$`Event Name`))
```

There are `r n_spp` unique species in this dataset.  It contains `r nrow(biol_records)` individual species observations, accross `r n_events` unique events in `r n_rivers` rivers.  
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
plot(table(as.numeric(biol_records$Year)), ylab = "number of records", xlab = "year")
```

##examine new data

compare it to the old ers_data2
```{r}
# cbind(names(biol_records), names(ers_data2))
# 
# #col 12 differs: [12,] "Event Name"              "NE Full Sample Name"    
# olddata <- ers_data2 #[-12]
# newdata <- biol_records #[-12]
# names(olddata) <- make.names(names(newdata))
# names(newdata) <- make.names(names(newdata))
# cbind(names(newdata), names(olddata))
# 
# 
# library(compareDF)
# 
# compareDF::compare_df(newdata, olddata, group_col = "Event.Name")
```

```{r}

```


### start import
#Full sSample data

#### Get a handle on numeric data

Try to find all the numeric values
```{r}
#new column, convert abundance to numeric. 
biol_records$abund_num <- as.numeric(biol_records$Abundance_Data)
skimr::skim(biol_records$abund_num)
#all data containing a non-numeric is now NA, all numeric is numeric
plot(biol_records$abund_num)
hist(biol_records$abund_num, breaks = 100)
hist(biol_records$abund_num, breaks = 1000, xlim = c(0,100))
```


```{r}
#split numeric and character abundance values: 
#create numeric column
abund_num <- biol_records %>% 
  select(abund_num) %>%  
  filter(is.na(abund_num) == FALSE) 
abund_num <- abund_num$abund_num

n_distinct(abund_num)
length(abund_num)
```

```{r}
ers_hist <- hist(abund_num, breaks = 100, plot = FALSE)
knitr::kable(tibble(ers_hist$breaks[-1], ers_hist$counts))
```
The vast majority of abundance values aer below 20.  

#### Get a handle on non-numeric data

```{r}
#create a character abundance column
biol_records <- add_column(biol_records, abund_char = biol_records$Abundance_Data) 
  
biol_records$abund_char <- replace(biol_records$abund_char, 
                               is.na(biol_records$abund_num)==FALSE, 
                               NA)
  
  
```




```{r}
n_distinct(biol_records$abund_char)
length(biol_records$abund_char[which(is.na(biol_records$abund_char)==FALSE)])
```
552 observations; 17 unique abundance values

- same as ers_data2


```{r}
ers_char <- data.frame(table(biol_records$abund_char))
ers_char <- arrange(ers_char, -Freq)
knitr::kable(ers_char)
```

##what about sample types
```{r}
knitr::kable(
  biol_records %>%
    count(Sample_Type) %>%
    arrange(-n)
)
```

```{r}
unique(biol_records$Sample_Type)
```


```{r}
table(tolower(biol_records$Sample_Type))

```


Clean up sample type coding
```{r}
biol_records <- biol_records %>% 
  mutate(Sample_Type_old = Sample_Type, 
         Sample_Type = recode_factor(tolower(biol_records$Sample_Type),
             `20 minute dig` = "excavation",
             `20 minute timed search` = "hand_search",
             `30 minute timed sample` = "unknown",
             aquatic = "aquatic",
             `casual record` = "ad-hoc",
             excavation = "excavation",
             `field observation` = "ad-hoc",
             `hand collection` = "hand_search",
             `hand search` = "hand_search",
             `hand search (25min)` = "hand_search",
             `hand searching` = "hand_search",
              handsearch = "hand_search",
             `handsearch/excavation` = "excavation",
             netting = "sweep",
             pitfall = "pitfall",
             `pitfall trap` = "pitfall",
             pitfalls = "pitfall",
             quadrat = "hand_search",
             splashing = "splashing",
             sweep = "sweep",
             `timed dig` = "excavation",
             `timed hand collection` = "hand_search",
             `timed hand search` = "hand_search",
             `water trap` = "water_trap"
             ))

table(tolower(biol_records$Sample_Type))

```



```{r}
table(tolower(biol_records$Sample_Type_old))
```



does data type relate to sample type
```{r}
sample_data_types <-
  biol_records %>%
    group_by(Sample_Type) %>%
    count("abund_character" = is.na(abund_char)==FALSE, 
        "abund_number" = is.na(abund_num)==FALSE) %>% 
    filter(abund_character == TRUE | abund_number == TRUE) 

sample_data_types$abund_character[which(sample_data_types$abund_character==TRUE)] <- 
  sample_data_types$n[which(sample_data_types$abund_character==TRUE)]
  
sample_data_types$abund_number[which(sample_data_types$abund_number==TRUE)] <- 
  sample_data_types$n[which(sample_data_types$abund_number==TRUE)]
  
#sample_data_types$n <- NULL

knitr::kable(sample_data_types)
```


##Fix sample timing
```{r}
table(biol_records$Sample_Type_old, biol_records$Sample_Duration)

```

```{r}
biol_records$Sample_Duration[which(biol_records$Sample_Type_old %in% c("20 minute dig", "20 minute timed search"))] <- "20 min"

biol_records$Sample_Duration[which(biol_records$Sample_Type_old %in% c("30 minute timed sample"))] <- "30 min"

biol_records <- biol_records %>% 
  mutate(Sample_Duration = recode_factor(Sample_Duration,
                                         `80-100`  = "80-100 min",
                                         `80-100 minutes` = "80-100 min"))




table(biol_records$Sample_Type_old, biol_records$Sample_Duration)
```



## Look at surveys


```{r}
# 
# names(ers_data2)
# 
# table(ers_data2$`NE Full Sample Name`)
```


#Sort out events

Note, I've renamed some of the columns for clarity: col A = 'Event',  Col C = number of pitfall replicates (P), col D = number of timed  search/excavation replicates(T), Col G = number of pitfalls per replicate.  


```{r}
biol_records$event <- biol_records$`Event Name` 

table(biol_records$event)
```


```{r}
events_code
```


```{r}
biol_events <- biol_records %>% select(event) %>% unique() %>% arrange(event) %>% 
  mutate(biol = "biol")

#%>% filter(!event %in% events_code$event)

events_code_lookup_apr_2019 <- full_join(events_code, biol_events) %>% 
  select(event, ers, sel, biol, event_code) %>% arrange(event)

#write_csv(events_code_lookup_apr_2019, "../data/events_code_lookup_apr_2019.csv")

#make changes in excel

events_code <- read_csv("../data/events_code_lookup_apr_2019.csv")

```


```{r}
#bind codes back into biol_records
biol_records <- left_join(
  biol_records, 
  select(events_code, event, event_code), 
  by = "event") %>% 
  mutate(event = as.factor(event)) %>% 
  mutate(event_code = as.factor(event_code))
```

```{r}
biol_records %>% group_by(event_code) %>% count(event_code)
```



##Sort out date field
```{r}
unique(biol_records$Date)
```


```{r}
# date validation rules: 
# - if span of dates, choose start date
# - if just a year, set to NA
# - if range of years, set to NA
# - NA, keep as NA

biol_records <- biol_records %>% 
  mutate(date_orig = Date) %>% 
  mutate(Date = str_replace_all(
    Date, 
    c(
    "1997" = NA,
    "1987" = NA,
    "12/06/2003- 23/06/2003" = "12/06/2003",
    "1998"  = NA,
    "NA" = NA,
    "11/06/2013 for two weeks" = "11/06/2013",
    "30/04/2003- 15/05/2003" = "30/04/2003",
    "26/04/2002- 09/05/2002" = "26/04/2002",
    "17/05/2003- 04/06/2003" = "17/05/2003",
    "13/06/2003- 23/06/2003" = "13/06/2003",
    "30/06/2002- 14/07/2002" = "30/06/2002",
    "19/09/2001- 03/10/2001" = "19/09/2001",
    "25/04/2002- 09/05/2002" = "25/04/2002",
    "1990 - 1996" = NA,
    "02 to 06/06/2015" = "02/06/2015",
    "18/09/2001- 03/10/2001" = "18/09/2001",
    "24/04/2002- 09/05/2002" = "24/04/2002",
    "13/06/2002- 14/07/2002" = "13/06/2002"))) %>% 
  droplevels()

unique(biol_records$Date)
```

```{r}
unique(biol_records$year_orig)

biol_records <- biol_records %>% 
  mutate(year_orig = Year) %>% 
  mutate(Year = replace(Year, is.na(Year), str_sub(Date, -4, -1))) %>%
  mutate(Year = replace(Year, is.na(Year), str_sub(Date, -4, -1)))  #weirdly I need too repeat this to get all the NAs

biol_records %>% filter(is.na(Year)) %>% select(Year, year_orig, Date)
```



## replication
```{r}
#create location replicates column
biol_records <- biol_records %>%
  mutate(smpshort = str_replace_all(
    Sample_Type, c(
      "hand_search" = "HS", "pitfall" = "PF", "excavation" = "EX", 
      "ad-hoc" = "AH", "sweep" = "SW", "aquatic" =  "AQ", 
      "unknown" = "UK", "water_trap" = "WT", "splashing" = "SP"))) %>% 
  mutate(locshort = str_remove_all(Location_Name, "[ ,.]")) %>% 
  mutate(repl_loc = paste(event_code, smpshort, locshort, sep = "-")) 

biol_records %>% group_by(event_code, Sample_Type, Location_Name) %>% 
  summarise(n.loc = length(unique(repl_loc))) %>% 
  arrange(-n.loc) 

biol_records %>% group_by(event_code, Sample_Type) %>% 
  summarise(n.loc = length(unique(repl_loc))) %>% 
  arrange(-n.loc) 
```


```{r}
#create date replicate column
biol_records <- biol_records %>%
  mutate(dateshort = str_remove_all(Date, "/")) %>%
  mutate(repl_date = paste(event_code, smpshort, dateshort, sep = "-"))

#create date-location replicates column
biol_records <- biol_records %>% 
  mutate(repl_dateloc = paste(event, smpshort, dateshort, locshort, sep = "-"))  

head(unique(biol_records$repl_dateloc))
  
biol_records %>% group_by(event, Sample_Type, Date, Location_Name) %>% count() %>% arrange(-n)

biol_records %>% group_by(event, Sample_Type) %>% summarise(n.dateloc = length(unique(repl_dateloc))) %>% 
  arrange(-n.dateloc) 
```



##Select field data
```{r}
field_data <- biol_records %>% 
  select(spp_name = Species_name, 
         abund_num,
         abund_char,          
         river = `Sample River`,
         event = event, 
         event_code = event_code,
         date = Date,
         year = Year, 
         sample_type = Sample_Type,
         location = Location_Name,
         repl_dateloc = repl_dateloc,
         repl_date = repl_date,
         replicate = repl_loc,
         sadler_bell = "Sadler & Bell method?",
         search_pitfall = "Handsearch & Pitfall",
         search_excavation = "Handserach & excavation",
         sample_duration = "Sample_Duration"
         )

field_data %>% 
  #group_by(event, sample_type) %>% 
  count(sadler_bell)
  
field_data %>% 
  #group_by(event, sample_type) %>% 
  count(sample_type, sadler_bell) %>% 
  arrange(-n)

field_data %>% 
  count(sample_type)

```

```{r}
#select field data
field_data_selected <- filter(field_data, sample_type %in% c("excavation", "hand_search", "pitfall")) %>% droplevels()

field_data_selected
```



```{r}
field_data_selected %>%
  summarise(n.events = n_distinct(event_code), n.rep_loc = n_distinct(replicate),
            n.rep_date = n_distinct(repl_date), n.rep_dateloc = n_distinct(repl_dateloc))
```

##Correct species names

refer to ripInverts-dataImport_3.Rmd

```{r}
#spp_matched_corrected <- read.csv("../data/speciesmatch_corrected.csv") 

spp_matched_corrected <- read.csv("../data/speciesmatch_further_correctionsApr2019.csv") 
```


```{r}
field_data_selected <- read.csv("../data/field_data_selected_Apr2019.csv")

#correct the names in the original data
tempnames <- names(field_data_selected)

temp_field_data_selected <- field_data_selected %>% 
  mutate(old_name = spp_name) %>% 
  left_join(spp_matched_corrected, by = "old_name") %>% 
  mutate(spp_name = match) %>%
#  filter(include == "y") %>% 
  select(old_name, tempnames)


```






```{r}
#species richness 
field_data_selected %>% 
  mutate(sample_type = recode(sample_type, 
                              "excavation" = "timed", 
                              "hand_search" = "timed")) %>% 
  group_by(event, sample_type) %>% 
  summarise(n_spp = n_distinct(spp_name),
            n_rec = length(spp_name))
```


##Check if new-data is alread incorporated
```{r}
raw <- read.csv("../data/2018 riparian update corrected.csv")
newdata <- raw
dim(newdata)
table(newdata$Grid.ref)
biol_records %>% filter(Sample_Grid_Ref %in% unique(newdata$Grid.ref)) %>%  dim()
```

hard to say 

##Select subset of events
```{r}
field_data_selected %>% select(year, event_code) %>% unique() %>% 
  arrange(-as.numeric(year)) %>% pull(event_code)
```



```{r}
event_code_selected <- c("Frome_2017_2", "Frome_2017_1", "Frome_2017_3", "Lugg_2014_1",
                         "Lugg_2014_3", "Lugg_2014_4", "Wye_2014_6", "Wye_2014_7", 
                         "Wye_2014_4", "Wye_2014_3", "Wye_2014_5", "Lugg_2014_2", 
                         "Wye_2014_1", "Dove_2013", "Till_2013", "Beamish_2013", 
                         "Wooler_2013")

field_data_extract <- field_data_selected %>% 
  filter(event_code %in% event_code_selected) %>% droplevels() 

field_data_extract %>% count(event_code)
```






#Export data
```{r}
# write_csv(field_data_selected, "../data/field_data_selected_Apr2019.csv")
# write_csv(field_data_extract, "../data/field_data_extract_Apr2019.csv")
```







